---
title: "3.1 Identify extreme event according specific constraints"
author: "Victoria Delannoy, Nicolas Loiseau, Sébastien Villéger"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{3.1 Identify extreme event according specific constraints}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Aims of the function *BEE.id.extreme_events()*
Identify the extreme event according to a set of constraints (e.g. minimum duration, merging close events).
*BEE.id.extreme_events()* transforms the extreme days identified with *BEE.id.extreme_day()* into coherent extreme events by applying temporal criteria such as minimum duration, merging events close in time, and removing isolated days.
The two functions are complementary: binarise detects when conditions are extreme, while true_event defines which sequences of extreme days constitute meaningful events.

# Usage
## Arguments:
**binarized_spatraster**: is a spatraster with one layer per day. Each layer covers the same area and time is continuous (no layers are missing). If you are not using a fixed threshold, the baseline spatraster computed using BEE.calc.baseline() must be computed using the same dataset as yourspatraster for comparison without bias.  
**n**: The minimum number of consecutive days above the baseline/threshol required for a series of days more extrem than the baseline to be considered as an extreme event (≥).  
**d**: The maximum number of days for which two series of values above the threshold/baseline, separated by some days below the threshold/baseline, can be considered a single extreme event (the maximum distance in days between the two events to be merged <=).  
**nbis**: The minimum number of days in a series of 1 to allow merging with
#' serie of 1 longer than *'n'* and distant from *'d'* days or fewer.
#' /!\ *'nbis'* must be inferior or equal to *'n'*. This avoid extending an extreme event for only a few extreme day. For instance: 0000100111110010000 will be considere as an extreme event
of duration 11 days when only using n = 5 and d= 2. However, if you set nbis=2, lonely extreme days will not be merged and the duration of the above event will be 5 days. It allows to be more or less precautionious on the temporal extent of the extreme event.


## Warnings:
No warning implemented yet.  
## Messages:
No messages implemented yet.

## Output
Returns a list with 2 elements.  
The first one is called **"binarized_spatraster_corrected"**. This SpatRaster contains the identified extreme events. Based on the criteria specified in the function arguments, extreme events are derived from the binarised extreme days by applying pixel-wise temporal rules such as minimum event duration, merging of events that are close in time, and the removal of isolated extreme days. These events are used in the next function of the pipeline.  
The second element is called **"extreme_event"**. This is a list with one data.table per pixel. Each data.table has a date for each row (and as many dates as there are layers in yourspatraster). 

The dataframes in the output list "extreme_event" contains the followings collumns :

|   Column name  |             Description            |   Class   |  Unit   |
|:--------------:|:----------------------------------:|:---------:|:-------:|
| pixel_id       | Pixel number in the spatraster     | integer   |         |
| original_value | Day more extreme than baseline = 1 | double    |         |
|                | Day less extreme than baseline = 0 |           |         |
| cleaned_value  | Day belong to an exteme event = 1  | double    |         |
|                | Day outside of an extreme event = 0|           |         |
| event_id       | Number of the extreme event in the | integer   |         | 
|                | time serie                         |           |         |
| duration       | Duration of the period (extreme    | integer   | days    |   
|                | event or in between 2 events)      |           |         |   
| date           | Date associated to the observation |double/date|         |  
| ID             | The event identifier \*            | character |         |

\* built using the pixel number, the first day of the event and the last day of the event and the event_id for that pixel. ID and event_id are proper to each pixel.  

If the position in the list corresponds to a pixel number for which the pixel value is consistently NA, the data frame is a null object.

# Note
The *BEE.calc.corrections()* function allows to check how many modifications your choices of settings in BEE.calc.tru_event have induced compare to row binerized data. Please check *3.bis. Check the number of modifications induced by the BEE.id.extreme_events() settings.*

# Examples
## Load data:
```{r setup}
library(BioExtremeEvent)
#Binarized spatraster built using BEE.id.extreme_day()
file_name_1 <- system.file(file.path("extdata", "copernicus_binarized.tiff"),
                                   package = "BioExtremeEvent")
binarized_data <- terra::rast(file_name_1)
```

## Apply filter:
```{r}
events <- BEE.id.extreme_events(binarized_spatraster = binarized_data, n = 5, d = 2)
```

## Spatraster output:
```{r}
# Get the corrected spatraster:
binarized_corrected_spatraster <- events[[1]]

# Plot a layer
date <- terra::time(binarized_corrected_spatraster[[800]])
terra::plot(binarized_corrected_spatraster[[800]],
            main = as.character(date))
# Here 850 refers to the layers of the spatraster <-> 800th days of the spatraster.
```

## Get the list of data.table with former and corrected values and with dates and duration:

```{r}
binarized_corrected_df <- events[[2]]
head(binarized_corrected_df[[80]])
# Here 80 refers to the pixel id. The 850th day correspond to the 850th row.
```


## How to save the ouputs:
Due to the different nature of the two outputs, they must be saved separately.

```{r, eval = FALSE}
## Save the spatraster
binarized_corrected_spatraster <- events[[1]]
terra::writeRaster(binarized_corrected_spatraster, "your_path/data/binarized_corrected_spatraster.tiff")

## Save the list of df
binarized_corrected_df <- events[[2]]
saveRDS(binarized_corrected_df, file = "your_path/data/binarized_corrected_df.rds")
```

Please save both because they are both used in the next steps of the pipeline.