---
title: 5. Calculate metrics on the shape of an extreme event."
author: "Victoria Delannoy, Nicolas Loiseau, Sébastien Villéger"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{5. Calculate metrics on the shape of an extreme event.}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Aims of the function *BEE.calc.metrics_morpho()*.
Compute daily metrics of spatial features of the extreme event (only in surface, 2D).

# Usage
## Arguments:
**corrected_rasters** :
  The SpatRaster you want to analyse (first output of 
  *BEE.id.extreme_events()*).
**start_date** :
  Allows to perform the analysis on a specific  time_frame, this allows to 
  save computation time.
**end_date** :
  Allows to perform the analysis on a specific  time_frame, this allows to 
  save computation time.
**per_pix** :
  Use TRUE if you want a list with one dt per pixel as an output.
**crs** :
  To get accurate length and area, data in longitude latitude must
  be converted into meters data that take in account earth sphericity, the
  conversion depend on studdied region, thus you must specify the one the most
  appropriated for your data. To find a suitable EPSG you can
  check on https://spatialreference.org/explorer.html or https://epsg.io/ or build your own EPSG here https://projectionwizard.org/.


## Warnings:
You will receive a warning if :

-   *"start_date"* and/or *"end_date"* formats do not match the format of the spatrasters layer name. Please note that some mismatches in the date format may still not be detected. To be sure, compare your *"start_date"* and *"end_sate"* with terra::names(corrected_rasters[[1:10]]).

## Messages:
You will receive a message in the following case :  
-    The number of NA pixels is not consitent through the layers, some metrics

## Output

You will get a list of dataframes contains the informations in the table bellow and a spartaster with all the extreme events, identified daily:
*Below, 'patch' is used to describe a group of neighbouring pixels on a given date.*

|   Column name        |             Description              |   Class   |  Unit          |
|:--------------------:|:------------------------------------:|:---------:|:--------------:|
| pixel_id             | Pixel number in the spatraster       | integer   |                | 
| x                    | Longitude                            | double    | decimal degree |
| y                    | Latitude                             | double    | decimal degree |
| patch_id             | Extreme event polygon identifier     | double    |                |
| cell_size            | Area of the pixel                    |           | m²             |
| boredering           | Whether the pixel is on the border of| double    | 1 or 0         |
|                      | the patch (1) or not (0).            |           |                |
| patch_area           | Total area of the extreme event on a | double    | m²             |
| core_area            | Area of non-boredering pixels        | double    | m²             |
|                      | of an extreme event on a given day   |           |                | 
| core_area_index      | patch_are/core_area                  | double    | no unit        |
| n_pixel              | number of pixels belong to the       | double    | pixels         |
|                      | extreme event                        |           |                |
| pixel_EE_tot         | Total number of pixels in extreme    | double    | pixels         |
|                      | event (all extreme events that       |           |                |
|                      | occured on that day)                 |           |                |
| cover_percent        | Percentage of non-NA area that is    | double    | percentage     |
|                      | covered by this extreme event (cf ID)|           |                |
| n_core_pixel         | Number of pixels in the extreme event| double    | pixels         |
|                      | that do not border the patch         |           |                |
| date                 | Date associated to the observation   | Date      |                |
| ID                   | The event identifier \*              | character |                |
| perimeter            | Perimeter associated to the extreme  | double    | m              |
|                      | event patch                          |           |                |
| centroid_x           | centroid longitude of the patch      | double    | degrees        |
| centroid_y           | centroid latitude of the patch       | double    | degrees        |
| perim_pixel_ratio    | Pixels on patch limit / total number |           | no unit        |
|                      | pixels in the patch                  |           |                |
| perim_area_ratio     | Perimter (in m) / area (in m² )      | double    | m⁻¹           |
| shape_index          | 0.25*perimeter/sqrt(area)            | double    | no unit        |
| fractal_cor_index    | 2*log(0.25 *perimeter)/log(area)     | double    | no unit        |
| contiguity_index     | Number of pixel from a patch within  | double    | pixels         |
|                      | the 8 pixels around each pixel from  |           |                |
|                      | the patch. For more info see:        |           |                |
|                      | ?landscapemetrics::lsm_p_contig      |           |                |
| ell_dir              | The angle between the north direction| double    | degrees        |
|                      | and the longest axis of the smallest |           |                | 
|                      | ellipse that surrounds a patch.      |           |                |
| ell_crop_area_m2     | Ellipse area within the raster       | double    | no unit        |
| ell_tot_area_m2      | Ellipse total area                   | double    | m²             |
| ell_long_axis        | Longest ellipse axis length          | double    | m              |
| ell_short_axis       | Shortest ellipse axis length         | double    | m              |
| patch_ell_ratio      | Patch area / ellipse area            | double    | no unit        |

\* Same format as the in the spatraster provided (here yyyy-mm-dd)  

## Find out more about ellipse computation and related metrics:

The ellipses computed in this package are the smallest ellipses that encapsulate the centres of all the pixels within a patch. They are computed daily. Since the calculation is based on the coordinates (in degrees) of the centre of the pixels in a patch, no ellipse can be computed for patches that form a line (vertical, horizontal or diagonal). Indeed, since all the coordinates used to compute the ellipse are aligned, no width can be assessed.
Ellipse computations are performed in two steps: first, the coordinates (in degrees) are computed based on the centre pixel coordinates (in degrees) using the cluster::ellipsoidhull() function. The axis lengths, azimuth or area computed by cluster::ellipsoidhull() cannot be used directly because cluster::ellipsoidhull() does not take into account deformation due to the degree reference. To avoid this issue, the ellipse properties are re-computed 'manually' using the coordinates of the ellipse perimeter (in degrees), and the axis lengths are calculated using functions such as terra::expanse(). Functions in the terra package compute distances and areas using projection onto a sphere, which is much more precise than projection onto a plane. According terra : "For vector data, the best way to compute area is to use the longitude/latitude CRS. This is contrary to (erroneous, but popular) belief that you should use a planar coordinate reference system." 
Please do not hestitate to check the function code available on the [github page](https://github.com/VicoDelannoy/BioExtremeEvent/blob/main/R/bee_calc_metrics_morpho.R) starting line 369.

# Examples
## Load data:
We need the spatraster "stacked_rasters_corrected", wich is an ouptut of BEE.calc.id_extreme_event().
```{r setup}
library(BioExtremeEvent)

file_name_1 <- system.file(file.path("extdata", 
                                      "binarized_corrected_spatraster.tiff"),
                                      package = "BioExtremeEvent")
copernicus_data_celsius <- terra::rast(file_name_1)
```

## Get metrics by pixel:

```{r}
# Warning, this can takes a minute or two.
list_morpho_metrics <- BEE.calc.metrics_morpho(copernicus_data_celsius, start_date = "2024-05-01", end_date = "2024-11-30", per_pix=TRUE) 
```

## Get metrics by patch:
```{r}
list_morpho_metrics2 <- BEE.calc.metrics_morpho(copernicus_data_celsius, start_date = "2024-05-01", end_date = "2024-11-30", per_pix=FALSE)
```


## How to save the ouputs:
If you used per_pix=TRUE, outputs are a list of dataframe, with one dataframe per pixel and a spatraster. 
To save your list of dataframe:
```{r, eval = FALSE}
df_list <- list_morpho_metrics[[1]]
saveRDS(df_list, file = "your_path/data/metrics_morpho_pixel.rds")
```

To save the spatraster:

```{r, eval = FALSE}
patch_spatraster <- list_morpho_metrics[[2]]
terra::writeRaster(patch_spatraster, "your_path/data/patch_spatraster.tiff")
```

I you used per_pix=FALSE the ouptuts are a dataframe (within a list) and a spataraster. Spatraster do not differ whether you use per_pix =TRUE or per_pix =FALSE.
To save the dataframe with morphometrics values per patch :

```{r, eval = FALSE}
patch_df <- list_morpho_metrics1[[1]] # get first object of the ouput
saveRDS(patch_df, file = "your_path/data/metrics_morpho_patch.rds")

# To avoid saving the dataframe as a list of one object, you can extract it from the list before saving:
patch_df_unlisted <- patch_df[[1]] #extract the df from the list
write.csv(patch_df_unlisted, file = "your_path/data/metrics_morpho_patch.csv")
```


# Plots

## Plot:
Quick plot of patches for a given day
```{r}


```

## Plot:
Plot a metric throught time for a given position (e.g. a given pixel):
```{r, fig.width = 20, fig.height =5, warning = FALSE}

```

