---
title: 5. Calculate metrics on the shape of an extreme event."
author: "Victoria Delannoy, Nicolas Loiseau, Sébastien Villéger"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{5. Calculate metrics on the shape of an extreme event.}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Aims of the function *BEE.calc.metrics_morpho()*.
Compute daily metrics of spatial features of the extreme event (only in surface, 2D).

# Usage
## Arguments:
**extreme_event_spatraster** :
  The SpatRaster you want to analyse (first output of 
  *BEE.id.extreme_events()*).
**start_date** :
  Allows to perform the analysis on a specific  time_frame, this allows to 
  save computation time.
**end_date** :
  Allows to perform the analysis on a specific  time_frame, this allows to 
  save computation time.
**per_pix** :
  Use TRUE if you want a list with one dt per pixel as an output.  Use TRUE to be able to use the output in BEE.data.merge().



## Warnings:
You will receive a warning if :

-   *"start_date"* and/or *"end_date"* formats do not match the format of the spatrasters layer name. Please note that some mismatches in the date format may still not be detected. To be sure, compare your *"start_date"* and *"end_sate"* with terra::names(extreme_event_spatraster[[1:10]]).

## Messages:
You will receive a message in the following case :  
-    The number of NA pixels is not consitent through the layers, some metrics

## Output

You will get a list of dataframes containning the informations in the table bellow and a spartaster with all the extreme events, identified daily:
*Below, 'patch' is used to describe a group of neighbouring pixels on a given date.*

|   Column name        |             Description              |   Class   |  Unit          |
|:--------------------:|:------------------------------------:|:---------:|:--------------:|
| pixel_id             | Pixel number in the spatraster       | integer   |                | 
| x                    | Longitude                            | double    | decimal degree |
| y                    | Latitude                             | double    | decimal degree |
| patch_id             | Extreme event polygon identifier     | double    |                |
| cell_size            | Area of the pixel                    |           | m²             |
| boredering           | Whether the pixel is on the border of| double    | 1 or 0         |
|                      | the patch (1) or not (0).            |           |                |
| patch_area           | Total area of the extreme event on a | double    | m²             |
| core_area            | Area of non-boredering pixels        | double    | m²             |
|                      | of an extreme event on a given day   |           |                | 
| core_area_index      | patch_are/core_area                  | double    | no unit        |
| n_pixel              | number of pixels belong to the       | double    | pixels         |
|                      | extreme event                        |           |                |
| pixel_EE_tot         | Total number of pixels in extreme    | double    | pixels         |
|                      | event (all extreme events that       |           |                |
|                      | occured on that day)                 |           |                |
| cover_percent        | Percentage of non-NA area that is    | double    | percentage     |
|                      | covered by this extreme event (cf ID)|           |                |
| n_core_pixel         | Number of pixels in the extreme event| double    | pixels         |
|                      | that do not border the patch         |           |                |
| date                 | Date associated to the observation   | Date      |                |
| ID                   | The event identifier \*              | character |                |
| perimeter            | Perimeter associated to the extreme  | double    | m              |
|                      | event patch                          |           |                |
| centroid_x           | centroid longitude of the patch      | double    | degrees        |
| centroid_y           | centroid latitude of the patch       | double    | degrees        |
| perim_pixel_ratio    | Pixels on patch limit / total number |           | no unit        |
|                      | pixels in the patch                  |           |                |
| perim_area_ratio     | Perimter (in m) / area (in m² )      | double    | m⁻¹           |
| shape_index          | 0.25*perimeter/sqrt(area)            | double    | no unit        |
| fractal_cor_index    | 2*log(0.25 *perimeter)/log(area)     | double    | no unit        |
| contiguity_index     | Number of pixel from a patch within  | double    | pixels         |
|                      | the 8 pixels around each pixel from  |           |                |
|                      | the patch. For more info see:        |           |                |
|                      | ?landscapemetrics::lsm_p_contig      |           |                |
| ell_dir              | The angle between the north direction| double    | degrees        |
|                      | and the longest axis of the smallest |           |                | 
|                      | ellipse that surrounds a patch.      |           |                |
| ell_crop_area_m2     | Ellipse area within the raster       | double    | no unit        |
| ell_tot_area_m2      | Ellipse total area                   | double    | m²             |
| ell_long_axis        | Longest ellipse axis length          | double    | m              |
| ell_short_axis       | Shortest ellipse axis length         | double    | m              |
| patch_ell_ratio      | Patch area / ellipse area            | double    | no unit        |

\* Same format as the in the spatraster provided (here yyyy-mm-dd)  

## Find out more about ellipse computation and related metrics:

The ellipses computed in this package are the smallest ellipses that encapsulate the centres of all the pixels within a patch. They are computed daily. Since the calculation is based on the coordinates (in degrees) of the centre of the pixels in a patch, no ellipse can be computed for patches that form a line (vertical, horizontal or diagonal). Indeed, since all the coordinates used to compute the ellipse are aligned, no width can be assessed.
Ellipse computations are performed in two steps: first, the coordinates (in degrees) are computed based on the centre pixel coordinates (in degrees) using the cluster::ellipsoidhull() function. The axis lengths, azimuth or area computed by cluster::ellipsoidhull() cannot be used directly because cluster::ellipsoidhull() does not take into account deformation due to the degree reference. To avoid this issue, the ellipse properties are re-computed 'manually' using the coordinates of the ellipse perimeter (in degrees), and the axis lengths are calculated using functions such as terra::expanse(). Functions in the terra package compute distances and areas using projection onto a sphere, which is much more precise than projection onto a plane. According terra : "For vector data, the best way to compute area is to use the longitude/latitude CRS. This is contrary to (erroneous, but popular) belief that you should use a planar coordinate reference system." 
Please do not hestitate to check the function code available on the [github page](https://github.com/VicoDelannoy/BioExtremeEvent/blob/main/R/bee_calc_metrics_morpho.R) starting line 369.

# Examples
## Load data:
We need the spatraster "stacked_rasters_corrected", wich is an ouptut of BEE.calc.id_extreme_event().
```{r setup}
library(BioExtremeEvent)

file_name_1 <- system.file(file.path("extdata", 
                                      "binarized_corrected_spatraster.tiff"),
                                      package = "BioExtremeEvent")
extreme_event_spatraster <- terra::rast(file_name_1)
```

## Get metrics by pixel:

```{r}
# Warning, this can takes a minute or two.
list_morpho_metrics <- BEE.calc.metrics_morpho(extreme_event_spatraster, start_date = "2024-05-01", end_date = "2024-11-30", per_pix=TRUE) 
```

/!\ In the list of dataframe (first output of BEE.calc.metrics_morpho()) are NAMED according pixel_id, which allows you to extract a dataframe by it's NAME. If you extract a dataframe by its position in the list, you will not get the dataframe associed to the expected pixel.

```{r}
# If looking for the datafrmae associated to pixel number 214, use "214", not 214:
df_list <- list_morpho_metrics[[1]] # get the dataframe list
# Get dataframe associated to pixel 214:
pixel_214 <- df_list[["214"]] # Don't forget the "" !

# Get the 214^th dataframe :
pixel_unknown <- df_list[[214]]
pixel_unknown$pixel_id[1] ; names(df_list[214])
# The dataframe pixel_unkown is associated to pixel number 302
# This is why it is important to extract dataset from the list using the NAME (using "pixel number").
```

## Get metrics by patch x day :
```{r}
list_morpho_metrics_EE <- BEE.calc.metrics_morpho(extreme_event_spatraster, start_date = "2024-05-01", end_date = "2024-11-30", per_pix=FALSE)
```


## How to save the ouputs:
If you used per_pix=TRUE, outputs are a list of dataframe, with one dataframe per pixel and a spatraster. 
To save your list of dataframe:
```{r, eval = FALSE}
df_list <- list_morpho_metrics[[1]]
saveRDS(df_list, file = "your_path/data/metrics_morpho_pixel.rds")
```

To save the spatraster:

```{r, eval = FALSE}
patch_spatraster <- list_morpho_metrics[[2]]
terra::writeRaster(patch_spatraster, "your_path/data/patch_spatraster.tiff")
```

I you used per_pix=FALSE the ouptuts are a dataframe (within a list) and a spataraster. Spatraster do not differ whether you use per_pix =TRUE or per_pix =FALSE.
To save the dataframe with morphometrics values per patch :

```{r, eval = FALSE}
patch_df <- list_morpho_metrics_EE[[1]] # get first object of the ouput
saveRDS(patch_df, file = "your_path/data/metrics_morpho_patch.rds")

# To avoid saving the dataframe as a list of one object, you can extract it from the list before saving:
patch_df_unlisted <- patch_df[[1]] #extract the df from the list
write.csv(patch_df_unlisted, file = "your_path/data/metrics_morpho_patch.csv")
```


# Plots

## Very basic plot of a metric through time:
Quick plot of patches for a given pixel.

```{r}
library(ggplot2)

# Metrics from a given position (pixel n°130):
df_list <- list_morpho_metrics[[1]]
pixel_302 <- df_list[["302"]]
# Convert area from m² to km²:
pixel_302$patch_area_km2 <- pixel_302$patch_area / 1e6

# Plot:
ggplot(pixel_302, aes(x = date,
                      y = patch_area_km2)) +
  geom_line(na.rm = FALSE, color = "darkorange") +
  scale_x_date(date_breaks = "1 month",
               date_labels = "%Y-%m") +
  labs(x = "Date",
       y = expression("Patch area (km"^2*")")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5))
```


If the area is small (as in our example), there is little intrest in plotting a metric for different pixel according to time. In deed, since sevral pixels belong to the same patch, they will often have the same value of area, perimeter etc. 

## Plotting patch area according azimut of the ellispe fitted on the patch:

```{r}
patch_df <- list_morpho_metrics_EE[[1]] #list of dataframe
patch_df <- dplyr::bind_rows(patch_df) #one big dataframe

#Converte m² to km²:
patch_df$patch_area_km2 <- patch_df$patch_area / 1e6

#Plot
library(ggplot2)

max_val <- max(patch_df$patch_area_km2, na.rm = TRUE)

ggplot(patch_df, aes(x = ell_dir, y = patch_area_km2)) +
  #on point per patch (per day):
  geom_point(color = "darkblue", size = 2) +
  coord_polar(theta = "x", start = pi, direction = 1) +
  scale_y_continuous(
    limits = c(0, max_val * 1.05),
    breaks = c(2000, 4000, 6000, 8000),
    expand = c(0, 0)
  ) +
    #dark lines for 4000 and 8000 (please ajust to the sacle of your data):
  geom_hline(yintercept = c(4000, 8000), color = "black", size = 0.5, linetype = "solid") +
    #grey dotted lines:
  geom_hline(yintercept = c(2000, 6000), color = "darkgrey", size = 0.5, linetype = "dotted") +
  labs(x = "Direction (°)", y = NULL, title = "patch area in km²") +
  theme_minimal() +
  theme(
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    plot.title = element_text(hjust = 0.5, vjust = -0.5, size = 11)
  ) +
    # add breaks values (circles):
  annotate("text", x = 0, y = 2000 + 300, label = "2000", color = "darkgrey", size = 3, hjust = 0.5) +
  annotate("text", x = 0, y = 4000 + 300, label = "4000", color = "black", size = 3, hjust = 0.5) +
  annotate("text", x = 0, y = 6000 + 300, label = "6000", color = "darkgrey", size = 3, hjust = 0.5) +
  annotate("text", x = 0, y = 8000 + 300, label = "8000", color = "black", size = 3, hjust = 0.5)
```
