---
title: "6. Calculate shortest distance (and direction) to escape from an extrem event."
author: "Victoria Delannoy, Nicolas Loiseau, Sébastien Villéger"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{6. Calculate shortest distance (and direction) to escape from an extrem event.}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Aims of the function *BEE.calc.escape()*.
Compute daily metrics related to the of the nearest "refuge area", its direction or position. The function only seek for refuge in 2D.

# Usage
## Arguments:
**extreme_events_output**: If a the full output of BEE.calc.id.extreme_events(), which means a list with 2 objects : the spatraster with binarized value AND the list of dataframe with one dataframe per pixel. See vignette *3.1 Identify extreme event according specific constraints* for more informations.

**start_date**: First day on which you want to start computing metrics. Optional argument. 

**end_date**: Last day on which you want to compute metrics. Optional argument.

**pixel**: A vector with x,y coordinates or a df of x,y coordinates or "all" if you want to compute distance metrics for all pixels in the raster provided through *"extreme_event_list"*.

**only_days_EE**: TRUE if you want an output including all days (even if they don't belong to an extreme event), FLASE if you want an output with only the date that belong to an extreme event.

**group_by_event** If TRUE, you will get a mean value per metrics per extreme event for every pixel analysed. If FALSE you will get daily value for each metrics. Use FALSE to be able to use the output in BEE.data.merge().
 

## Warnings:
You will receive a warning if :

-   The two elements of *"extreme_events_output"* argument don't cover the same time period. If you enccounter this issue, please make sure you are using the full output of *BEE.id.extreme_event()* without intermediate modifications.
-   The format of *'pixel'* argument is not recognized by the function. Please make sure that the provided dataframe contains only two columns, named x and y.
-   The spatraster provided through *"extreme_events_output"* have no pixel value equal to 0 or 1, it is very likely that it is not binarized. PLease check function *BEE.id.extreme_day()* and *BEE.id.extreme_event* or see vignettes n°2 and n°3.1 .

## Messages:
You will receive a message in the following case :  
-    Several pixels are "the closest pixel" which means there are sevral refuge area at the same distance. The azimut to the closest refuge pixel is the azimut to the pixel that as the lowest pixel id.
-   No extreme event was found on the given pixel over the time period provided through *"start_date"* and *"end_date"*.
-   For a given date, the raster (area provided) was fully covered by an extreme event, which means that it was not possible to identify a place to escape from the extreme event.
-   The combination of arguments is not handled by the function.

## Output

You will get a list of dataframes containning the informations in the table bellow:  


|   Column name        |             Description              |   Class   |  Unit          |
|:--------------------:|:------------------------------------:|:---------:|:--------------:|
| pixel_id             | Pixel number in the spatraster       | integer   |                |
| date                 | Date associated to the observation \*| Date      |                | 
| x                    | Longitude                            | double    | decimal degree |
| y                    | Latitude                             | double    | decimal degree |
| to_x                 | Longitude of the closest pixel       | character | decimal degree |
|                      | outside the extreme event.           |           |                |
| to_y                 | Latitude of the closest pixel        | character | decimal degree |
|                      | outside the extreme event.           |           |                |
| pixel_to_id          | Pixel number of the closest pixel    | character |                |
|                      | outside of the extreme event.        | integer   |                |
| distance             | Distance to the closest pixel outside| character | m              |
|                      | the extreme event                    |           |                |
| azimut               | Angle between north and the axis     | character | degrees        |
|                      | formed by the pixel to flee and the  |           |                |
|                      | refuge pixel                         |           |                |
| ID                   | The event identifier                 | character |                |
\* Same format as the in the spatraster provided (here yyyy-mm-dd)  

# Examples
## Load data:
We need the full output of *BEE.calc.id_extreme_event()..
```{r setup}
library(BioExtremeEvent)
# reconstruct the output of BEE.id.extreme_event() from data saved in local:
file_name_1 <- system.file(file.path("extdata", 
                                      "binarized_corrected_spatraster.tiff"),
                                      package = "BioExtremeEvent")
extreme_event_spatraster <- terra::rast(file_name_1)
file_name_2 <- system.file(file.path("extdata", 
                                      "binarized_corrected_df.rds"),
                                      package = "BioExtremeEvent")
extreme_event_list <- readRDS(file_name_2)

extreme_event <- list(extreme_event_spatraster,extreme_event_list)

# Build a DATAFRAME with position to studdy:
 GPS <- as.data.frame(terra::xyFromCell(extreme_event_spatraster[[1]], 
                          cell = terra::cells(extreme_event_spatraster[[1]])))
# This take in account every pixel in the raster, but you can also build your set of coordinates:
 GPS_2 <- data.frame(x = c(3.6883,4.4268), # Sète, Saintes-Marie-de-la-Mer
                     y = c(43.3786,43.4279))
```

## Get metrics by pixel:
```{r}
library(BioExtremeEvent)

escape <- BEE.calc.escape(
  extreme_events_output = extreme_event,
  start_date = "2023-01-01",
  end_date = "2025-12-15",
  pixel= GPS,
  only_days_EE = FALSE,
  group_by_event = FALSE
) 
```

## Get metrics per extreme events only:
```{r, message=FALSE}
library(BioExtremeEvent)

escape_EE_ONLY <- BEE.calc.escape(
  extreme_events_output = extreme_event,
  start_date = "2024-05-01",
  end_date = "2024-11-30",
  pixel= GPS,
  only_days_EE = TRUE,
  group_by_event = TRUE
) 
```

Note: if you use *"only_days_EE"* = FALSE and *"group_by_event"* = TRUE, you will get a line for the period in bteween extreme events, this allow to keep a continuous time line.

# How to save the ouputs:
If you used group_by_event=FALSE, output is a list of dataframe, with one dataframe per pixel. 
To save your list of dataframe:
```{r, eval = FALSE}
saveRDS(escape, file = "your_path/data/escape_pixel.rds")
```

If you used group_by_event=TRUE, output is a dataframe. 
```{r, eval = FALSE}
write.csv(escape_EE_ONLY, file = "your_path/data/escape_EE.csv")
```


# Plots

## Plot refugees around a pixel.

```{r}
# Extract the dataframe associated to pixel 108 from th list of df:
escape_108 <- escape[["108"]] 

# Plot:
library(ggplot2)
## Create circles to add on plot:
cercles <- seq(0, 50000, by = 10000) 
ggplot(escape_108, ggplot2::aes(x = as.numeric(azimut), y = as.numeric(distance))) +
  # Densité de points par saison
  stat_density_2d_filled(
    geom = "polygon",
    contour_var = "ndensity",
    show.legend = TRUE,
    alpha = 1
  ) +
  geom_point(data = data.frame(x = 0, y = 0),
             ggplot2::aes(x = x, y = y),
             shape = 4, size = 1, stroke = 0.5, color = "black", inherit.aes = FALSE) +
  # Dessiner les cercles concentriques blancs
  geom_hline(yintercept = cercles, color = "lightpink", linetype = "solid", alpha = 0.3) +
  coord_polar(theta = "x", start = 0) +
  scale_x_continuous(
    limits = c(0, 360),
    breaks = seq(0, 360, by = 30),
    expand = c(0, 0)
  ) +
  theme_classic() +
  theme(
    strip.background = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.title = element_blank(),
    panel.grid.minor = element_blank()
  ) +
  ggplot2::labs(title = "Positions of daily closest refuge pixel from pixel 108", fill = "Density")

```

