---
title: "7. Merge the the ouput of BEE.calc.'..' functions in one dataframe."
author: "Victoria Delannoy, Nicolas Loiseau, Sébastien Villéger"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{7. Merge the metrics on extreme events in one dataframe.}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Aims of the function *BEE.data.merge()*
Merge the ouputs of *BEE.calc.metrics_point()*, *BEE.calc.metrics_morpho()* and/or *BEE.calc.escape()* in one df per pixel. Among other thing, this allows to link daily patch to an extreme event.

# Usage
## Arguments:
**data_metrics_point**:  
 The list of dataframe provided in the ouput of *BEE.calc.metrics_point()*.
**data_metrics_morpho**:  
 The list of dataframe provided in the output of *BEE.calc.metrics_morpho()* (The first object, the second on is a spatraster not used here.)
**data_escape**:  
 The list of dataframe provided in the output of *BEE.calc.escape()*.
**crs**
 The CRS of the Spatraster that you have used so far to compute the baseline and identify the extreme event. The function does not modify the CRS of your coordinates. CRS is only used to build polygons and check that the datasets you have provided cover the same area.

## Warnings:
You will receive a warning if :

-   The CRS you provided is not recognised by 'terra' package. Use the 'terra::crs()' function on the spatial arrays you have provided to the *BEE.calc.baseline()* or *BEE.calc.metrics_morpho()* functions, for example.  
-   You didn't provided datasets to merge together or you provided only one dataset.
-   Some dates are missing in one of the dataset or the resolution is not daily. 
-   Some of the provided datasets do not share dates, so merging is not possible. Please compute the metrics for the same time period.
-   The provided datasets do not cover the same spatial area, so merging is not possible. Please compute the metrics for the same spatial raster (without cropping during the workflow).

## Messages:
You will receive a message in the following case :  
-    One dataset does not cover the same time period as the others. The merge will only be performed for the time periods that overlap in each dataset.  


## Output
Returns a list of dataframe with on dataframe per pixel. Only pixel/coordinates that are present in all the dataset provided will be present in the ouput.  
To get more informations on the metrics definition please check the tables from the vignettes n°4, 5 and 6.

# Examples
## Load data:
```{r setup}
library(BioExtremeEvent)

file_name_1 <- system.file(file.path("extdata", 
                                      "escape_pixel.rds"),
                                      package = "BioExtremeEvent")
escape_df_list <- readRDS(file_name_1)

file_name_2 <- system.file(file.path("extdata", 
                                      "metrics_morpho_pixel.rds"),
                                      package = "BioExtremeEvent")
morpho_df_list <- readRDS(file_name_2)

file_name_3 <- system.file(file.path("extdata", 
                                      "metrics_points_pixel.rds"),
                                      package = "BioExtremeEvent")
point_df_list <- readRDS(file_name_3)

```

## Apply function on three dataset:
```{r}
library(BioExtremeEvent)

merged_ouputs <- BEE.data.merge( data_metrics_point = point_df_list,
                                 data_metrics_morpho = morpho_df_list, 
                                 data_escape = escape_df_list, 
                                 crs= "EPSG:4326")
```

You can also provide only two data_... arguments to the function.

# How to save the ouputs:
The output is a list of dataframe, you can save it as follow:

```{r, eval = FALSE}
saveRDS(merged_ouputs, file = "your_path/data/merged_ouputs.rds")
```

# Plots
You can know work on the outputs of sevral functions :

```{r}
pixel_108 <- merged_ouputs[["108"]] #Dont forget the quote or you may get an unexpected pixel

#Check if the more intense day are also the one for which the distance to escape from the extreme event are bigger. 
library(ggplot2)

ggplot(pixel_108, aes(x = as.numeric(anomaly_qt), y = as.numeric(distance)/1000)) +
  geom_point(color = "steelblue", size = 3, alpha = 0.7) +
  labs(
    title = "Anomaly vs Distance",
    x = "Anomaly (°C)",
    y = "Distance (km)"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 11),
    panel.grid.major = element_line(color = "gray90"),
    panel.grid.minor = element_blank()
  )
```

Since *BEE.data.merge()* only works with continuous daily data, the output also has continuous daily data. Therefore, many days in the output do not correspond to an extreme event.
You can remove them as follows:

```{r}
pixel_108 <- merged_ouputs[["108"]]
pixel_108_EE <- pixel_108[pixel_108$cleaned_value==1,]

library(ggplot2)

library(ggplot2)

ggplot(pixel_108_EE, aes(x = as.numeric(anomaly_qt), y = as.numeric(distance)/1000, color = patch_ell_ratio)) +
  geom_point(size = 3, alpha = 0.7) +
  geom_vline(xintercept = 0, color = "red", linetype = "solid", linewidth = 1) +
  annotate(
    "text",
    x = -0.5,
    y = Inf,
    label = "Not an extreme day but\nbelongs to an\nextreme event",
    color = "red",
    hjust = 1,
    vjust = 1.2,
    size = 5,
    fontface = "plain"
  ) +
  labs(
    title = "Anomaly vs Distance",
    x = "Anomaly (°C)",
    y = "Distance (km)",
    color = "patch_ell_ratio"
  ) +
  xlim(NA, 3) +
  scale_color_viridis_c() +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 11),
    panel.grid.major = element_line(color = "gray90"),
    panel.grid.minor = element_blank(),
    legend.position = "right"
  )

```