---
title: "BEE.calc.true_event"
author: "Victoria"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---

---
title: "BEE.calc.true_event"
author: "Victoria"
format: html
editor: visual
---

This function apply a filter to withdraw isolated events. For instance, is you consider that an event that last only 2 days should not be accounted for in future analysis, you can correct those value by 0. Here are the different filter you can apply :

Simple filter : HERE WE COULD INCLUDE AN INTERACTIVE INTERFACE TO TEST DIFFERENT PARAMETERS VALUES

Low complexity :

-   Minimum number of consecutive day above threshold to consider that there is an extreme event : *n* (\>=)

-   Maximum number of days bellow thershold to consider two series of value above threshold as a single extreme event (distance max in days btw 2 event to merge them): *d* (\<=)

Examples : *n*\>=5 *d*\>=2

0 0 0 1 1 1 1 **1** 0 0 0 = 1 event, but 0 0 0 1 1 1 1 0 0 0 = no event ;

0 0 0 1 1 1 1 1 0 0 1 1 1 0 0 0 = 1 event of 5 days, but 0 0 0 1 1 1 1 1 0 0 1 1 1 **1 1** 0 0 0 = 1 event of 12 days, but 0 0 0 1 1 1 1 1 0 0 **0** 1 1 1 1 1 0 0 0 = 2 instincts event of 5 days each.

Medium complexity (Hobday et al. 2016,2018) :

-   Minimum number of days distant from *d* days or less to an event of *n* days : *nbis*.

Example : *n* = 5 ; d = 2 ; nbis = 3

This is for the case you want to consider 0 0 0 1 1 1 1 1 0 0 1 1 **1** 0 0 0 as one event of 10 days but 0 0 0 1 1 1 1 1 0 0 1 1 0 0 0 as one event of 5 days. In the first case, if there are not five consecutive days above threshold right before or after (d=2) the 3 consecutive days above threshold, the 3 consecutive '1' will not be considered as an event as *n* still fixed to 5.

High complexity :

-   Unstable window *w* between to event in which you allow a certain proportion *p* of days of *w* to be bellow threshold if this windows is bordered by series of *w \*p c*onsecutive days above threshold on each size. *w* and *p* replace *d*. If *w i*s an odd number, the minimum number of days above threshold within *w* will be the value rounded to the superior round number. For instance if *w* = 7 and *p* = 0.3 a window of 7 days between to event of *n* consecutive days will not create two distinct event if at least 4 days are above threshold. ( *w* \* *p* = 7 \* 0.5 = 3.5 rounded to 4)

Examples : n = 5; w = 10; p = 0.5 -\> w \* p = 5

case that lead to a single big event of 15 days :

0 0 0 0 0 1 1 1 1 1 0 1 1 1 0 0 0 1 1 0 1 1 1 1 1 0 0 0 0 0

0 0 0 0 0 1 1 1 1 1 0 0 1 1 1 1 1 0 0 0 0 0 0 0 0 0 0 0 0 0 (you can also use low complexity *n* = 5 and *d* = 3)

0 0 0 0 0 1 1 1 1 1 0 1 0 1 0 1 0 1 0 1 1 1 1 1 1 0 0 0 0 0

case that lead to 2 distinct events :

0 0 0 0 0 1 1 1 1 1 0 1 1 0 0 1 0 1 0 0 1 1 1 1 1 0 0 0 0 0

0 0 0 0 0 1 1 1 1 1 0 0 0 1 1 1 1 0 0 0 1 1 1 1 1 0 0 0 0 0

0 0 0 0 0 1 1 1 1 1 0 1 1 1 1 0 0 0 0 0 1 1 1 1 1 0 0 0 0 0 (if you want to detect 1 event of 10 days then a distinct event of 5 days instead of 2 events of 5 days only you can use : low complexity with *n* = 4 and *d* = 1)

```{r}
BEE.calc.true_event <- function(binarized_EE, n, d = NULL, nbis = NULL, w = NULL, p = NULL){ 
    # Extract all rasters (If we have to think to a faster way we could filter so pixel that are always NA (land) are not processed and we could create "all_dates" starting from scratch, only knowing first day and last day and assuming there is no missing day) 
  all_data <- lapply(seq_along(binarized_EE), function(i) {
    j_indices <- seq(1, length(names(binarized_EE[[i]])),1)
    list(
      dates = names(binarized_EE[[i]][j_indices]),
      rasters = binarized_EE[[i]][[j_indices]]
    )
  })

  # Merge data to save memory
  all_dates <- unlist(lapply(all_data, `[[`, "dates"), recursive = FALSE)
  all_rasters <- rast(lapply(all_data, `[[`, "rasters"))
  # Clean memory
  rm(all_data)
  gc()
  
  start_date <- min(unlist(all_dates), na.rm = TRUE) # for later when I'll give the possibility to only work on a subset of the time période
  end_date <- max(all_dates, na.rm = TRUE)

  # Reorder informations in chronological order :
  ## get a list of the raster layers
  all_rasters <- lapply(1:nlyr(all_rasters), function(i) all_rasters[[i]])
  ## Transforme SpatRaster to list of rasters so it is easier to modify the order.
  sorted_indices <- order(all_rasters_names) 
  ## get a list of the current position of the rasters : the first number tell you where is currently the raster that should be first in chronological order
  sorted_rasters <- all_rasters[sorted_indices] #sort the rasters
  
  # Transform the list of sorted raster into a multilayers raster to be able to use terra:extract on cells
  stacked_rasters <- rast(sorted_rasters)
  pixel_time_series <- terra::extract(stacked_rasters, 1:ncell(stacked_rasters)) #one row per pixel, 1 column per time step
  # Transform to character the series of value for each pixels

  if (!is.null(n) & !is.null(d) & is.null(nbis) & is.null(w) & is.null(p)){ # code for low complexity

    correct_lowcomplexity_n_d <- function(pixel_time_series, c, n, d, pix){ # pixel_values <- pixel_time_series[1,] and pix=1; "pix" is a dynamic argument that is automatically indexed when calling the function, it identify each pixels (~ each rows of pixel_time_series)
      
       pixel_values <- as.numeric(pixel_time_series[c, ])
       rle_series <- rle(pixel_values) 
      # Clean isolated '1'
       One_to_0 <- which(rle_series$values == 1 & rle_series$lengths < n) #list isolated '1' (series < n ) by 0
       rle_series$values[One_to_0] <- 0 # replace isolated 1 by 0
       pixel_values_cleanned <- inverse.rle(rle_series) # values of each pixel trough time
       rle_series_cleanned <- rle(unlist(pixel_values_cleanned)) # re-identify the series of 0, this way series of zeros are re-defined to include 0 and 1 transformed to 0 in one same serie when they are next to each other.
       pixel <- data.table::data.table(Original_value = as.vector(pixel_values), #need to go through a dataframe because I don't know how to modify the ID directly in rle object
                           Cleanned_value = unlist(pixel_values_cleanned),
                           ID = rep(seq_along(rle_series_cleanned$lengths), rle_series_cleanned$lengths),
                           Nb_days = rep(rle_series_cleanned$lengths, rle_series_cleanned$lengths))
  
      # Rename event that should gathered because they are separated by d days (or less) bellow threshold
       zero_series <- unique(pixel$ID[(is.na(pixel$Cleanned_value) | pixel$Cleanned_value ==0) & pixel$Nb_days <= d]) #identify series of 0 or NA shorter than d (or equal to d) This imply that a sequence of NA shorter than d will not be considered as a limit of event
       zero_series <- setdiff(zero_series,c(1,max(pixel$ID))) # withdraw the beginning and end of the timeserie as it cannot be surrounded by '1' values
       to_modify <- sort(c(zero_series, zero_series + 1))  #list of futur index to rename zero series and one series under one same name when necessary, using the first one serie's name (only when 0 serie length <=d)
       modifier <- zero_series-1 #sequence of one before the short sequence of 0, <-> the sequence of one that will be merge with the next sequence of 1
       replacement_map <- setNames(rep(modifier, each = 2), to_modify) # associate the ID value to replace (firt line) with the value of replacement (second line)
       # Here above, if you have something like 111110111110111110 you will get aaabb as ID instead of aaaaa to adress this issue, I've added the lines bellow :
       pixel[, ID := ifelse(ID %in% names(replacement_map), as.numeric(replacement_map[as.character(ID)]), ID)]
       pixel[, Nb_days := rep(rle(ID)$lengths, rle(ID)$lengths)]
       pixel[, ID := paste0(pix, '_', ID)]
       return(pixel)
    }
    pixel_time_series <- as.data.table(pixel_time_series)
    gc()

    Event_corrected <- list()
    for (c in 1:nrow(pixel_time_series)){ # Here a loop works better than a mcapply because mcapply copies the data.table in each cores, which takes time and uses RAM.
      pix <- c
      Event_corrected[[c]] <- correct_lowcomplexity_n_d(pixel_time_series, c, n, d, pix)
    }
  }
  
  if (!is.null(n) & !is.null(d) & !is.null(nbis) & is.null(w) & is.null(p)){
    #here add code for medium complexity
  }
  if (!is.null(n) & is.null(d) & is.null(nbis) & !is.null(w) & !is.null(p)){
    #here add code for high complexity
  }
  gc()
  # Aply corrections to raster of 1 and 0
  stacked_rasters_corrected <- lapply(1:nlyr(stacked_rasters), function(t) { #stacked raster are rasters in the same order than in Event_corrected but before modifications
    correct_raster(stacked_rasters[[t]], t, Event_corrected)
  })

  stacked_rasters_corrected <- rast(stacked_rasters_corrected)
  names(stacked_rasters_corrected) <- sort(all_dates)

  gc()
  
  return(list(stacked_rasters_corrected = stacked_rasters_corrected, 
            Event_corrected = Event_corrected))
}
```

#function used in the big function :

```{r}
  # Aply corrections
  correct_raster <- function(raster, t, Event_corrected) {
    # get values of each layers
     corrected_values <- values(raster)
     Matching_cleanning_values <- sapply(Event_corrected, function(event) event$Cleanned_value[t])
    # Identify where corrections are needed /!\ 
    correction_indices <- which(!is.na(corrected_values) & !is.na(Matching_cleanning_values) & corrected_values != Matching_cleanning_values)
     # Apply corrections
     corrected_values[correction_indices] <- Matching_cleanning_values[correction_indices]
    # Modify raster layer values
    values(raster) <- corrected_values
    return(raster)
  }
```
